---
title: 'Stat Distributions '
author: "Braeden Kuether"
date: "7/23/2019"
output:
  html_document: default
  pdf_document: default
---

## Statistical Conditions

Using statistics from Basketball Reference. Player has played over 20 games
and over 6 minutes per game.

## Loading Packages

```{r load_packages, echo=T, message=FALSE, warning=FALSE, results="hide"}
lapply(
  c('heatmaply', # devtools::install_github('talgalili/heatmaply')
    'mclust', #install.packages('mclust')
    'dplyr', # devtools::install_github('hadley/dplyr')
    'plotly', # devtools::install_github('ropensci/plotly')
    'nbastatR', # devtools::install_github('abresler/nbastatR')
    'purrr', # devtools::install_github('hadley/purrr')
    'tidyr' # devtools::install_github('hadley/tidyr')
  ),
  library,
  character.only = T
)

options(digits = 4)

```

## Using statistics from Basketball Reference

```{r, message=FALSE, warning=FALSE, include=FALSE}

stat_2019 <- bref_players_stats(season = "2019", tables = c("advanced", "totals", "per game", "per 36"), include_all_nba = F, only_totals = TRUE, nest_data = FALSE, assign_to_environment = TRUE, widen_data = TRUE, join_data = TRUE, return_message = TRUE) %>%
   filter(countGames > 20) %>%
   filter(minutesPerGame > 6) %>%
   mutate(df_size = length(ptsPerGame) - 1) %>%
   mutate(ptsPerMinute = ptsPerGame / minutesPerGame, fgaPerMinute = fgaPerGame / minutesPerGame) %>%
   mutate(astPerMinute = astPerGame / minutesPerGame, tovPerMinute = tovPerGame / minutesPerGame) %>%
   mutate(trbPerMinute = trbPerGame / minutesPerGame, orbPerMinute = orbPerGame / minutesPerGame, drbPerMinute = drbPerGame / minutesPerGame) %>%
   mutate(blkPerMinute = blkPerGame / minutesPerGame, stlPerMinute = stlPerGame / minutesPerGame, pfPerMinute = pfPerGame / minutesPerGame)
  
size = length(stat_2019$ptsPerGame) - 1
```

## Points Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$ptsPerGame),]

#Mean
mean(stats_2019$ptsPerGame)
#Standard Deviation
sd(stats_2019$ptsPerGame)
#Spread
quantile(stats_2019$ptsPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$ptsPerGame, mean(stats_2019$ptsPerGame), sd(stats_2019$ptsPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = ptsPerGame), color = "red") +
  xlab("Player") +
  ylab("Points Per Game") +
  ggtitle("Visualizing Points Per Game Distribution")
```

## Points Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$ptsPerMinute),]

#Mean
mean(stats_2019$ptsPerMinute)
#Standard Deviation
sd(stats_2019$ptsPerMinute)
#Spread
quantile(stats_2019$ptsPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$ptsPerMinute, mean(stats_2019$ptsPerMinute), sd(stats_2019$ptsPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = ptsPerMinute), color = "red") +
  xlab("Player") +
  ylab("Points Per Minute") +
  ggtitle("Visualizing Points Per Minute Distribution")
```

## Assits Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$astPerGame),]

#Mean
mean(stats_2019$astPerGame)
#Standard Deviation
sd(stats_2019$astPerGame)
#Spread
quantile(stats_2019$astPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$astPerGame, mean(stats_2019$astPerGame), sd(stats_2019$astPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = astPerGame), color = "red") +
  xlab("Player") +
  ylab("Assists Per Game") +
  ggtitle("Visualizing Assists Per Game Distribution")
```

## Assists Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$astPerMinute),]

#Mean
mean(stats_2019$astPerMinute)
#Standard Deviation
sd(stats_2019$astPerMinute)
#Spread
quantile(stats_2019$astPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$astPerMinute, mean(stats_2019$astPerMinute), sd(stats_2019$astPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = astPerMinute), color = "red") +
  xlab("Player") +
  ylab("Assists Per Minute") +
  ggtitle("Visualizing Assists Per Minute Distribution")
```

## Total Rebounds Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$trbPerGame),]

#Mean
mean(stats_2019$trbPerGame)
#Standard Deviation
sd(stats_2019$trbPerGame)
#Spread
quantile(stats_2019$trbPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$trbPerGame, mean(stats_2019$trbPerGame), sd(stats_2019$trbPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = trbPerGame), color = "red") +
  xlab("Player") +
  ylab("Total Rebounds Per Game") +
  ggtitle("Visualizing Total Rebounds Per Game Distribution")
```

## Total Rebounds Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$trbPerMinute),]

#Mean
mean(stats_2019$trbPerMinute)
#Standard Deviation
sd(stats_2019$trbPerMinute)
#Spread
quantile(stats_2019$trbPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$trbPerMinute, mean(stats_2019$trbPerMinute), sd(stats_2019$trbPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = trbPerMinute), color = "red") +
  xlab("Player") +
  ylab("Rebounds Per Minute") +
  ggtitle("Visualizing Total Rebounds Per Minute Distribution")

```

## Offensive Rebounds Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$orbPerGame),]

#Mean
mean(stats_2019$orbPerGame)
#Standard Deviation
sd(stats_2019$orbPerGame)
#Spread
quantile(stats_2019$orbPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$orbPerGame, mean(stats_2019$orbPerGame), sd(stats_2019$orbPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = orbPerGame), color = "red") +
  xlab("Player") +
  ylab("Offensive Rebounds Per Game") +
  ggtitle("Visualizing Offensive Rebounds Per Game Distribution")
```

## Offensive Rebounds Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$orbPerMinute),]

#Mean
mean(stats_2019$orbPerMinute)
#Standard Deviation
sd(stats_2019$orbPerMinute)
#Spread
quantile(stats_2019$orbPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$orbPerMinute, mean(stats_2019$orbPerMinute), sd(stats_2019$orbPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = orbPerMinute), color = "red") +
  xlab("Player") +
  ylab("Offensive Rebounds Per Minute") +
  ggtitle("Visualizing Offensive Rebounds Per Minute Distribution")

```

## Defensive Rebounds Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$drbPerGame),]

#Mean
mean(stats_2019$drbPerGame)
#Standard Deviation
sd(stats_2019$drbPerGame)
#Spread
quantile(stats_2019$drbPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$drbPerGame, mean(stats_2019$drbPerGame), sd(stats_2019$drbPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = drbPerGame), color = "red") +
  xlab("Player") +
  ylab("Defensive Rebounds Per Game") +
  ggtitle("Visualizing Defensive Rebounds Per Game Distribution")
```

## Defensive Rebounds Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$drbPerMinute),]

#Mean
mean(stats_2019$drbPerMinute)
#Standard Deviation
sd(stats_2019$drbPerMinute)
#Spread
quantile(stats_2019$drbPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$drbPerMinute, mean(stats_2019$drbPerMinute), sd(stats_2019$drbPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = drbPerMinute), color = "red") +
  xlab("Player") +
  ylab("Defensive Rebounds Per Minute") +
  ggtitle("Visualizing Defensive Rebounds Per Minute Distribution")

```

## Steals Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$stlPerGame),]

#Mean
mean(stats_2019$stlPerGame)
#Standard Deviation
sd(stats_2019$stlPerGame)
#Spread
quantile(stats_2019$stlPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$stlPerGame, mean(stats_2019$stlPerGame), sd(stats_2019$stlPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = stlPerGame), color = "red") +
  xlab("Player") +
  ylab("Steals Per Game") +
  ggtitle("Visualizing Steals Per Game Distribution")
```

## Steals Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$stlPerMinute),]

#Mean
mean(stats_2019$stlPerMinute)
#Standard Deviation
sd(stats_2019$stlPerMinute)
#Spread
quantile(stats_2019$stlPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$stlPerMinute, mean(stats_2019$stlPerMinute), sd(stats_2019$stlPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = stlPerMinute), color = "red") +
  xlab("Player") +
  ylab("Steals Per Minute") +
  ggtitle("Visualizing Steals Per Minute Distribution")

```

## Blocks Per Game Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$blkPerGame),]

#Mean
mean(stats_2019$blkPerGame)
#Standard Deviation
sd(stats_2019$blkPerGame)
#Spread
quantile(stats_2019$blkPerGame)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$blkPerGame, mean(stats_2019$blkPerGame), sd(stats_2019$blkPerGame))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = blkPerGame), color = "red") +
  xlab("Player") +
  ylab("Blocks Per Game") +
  ggtitle("Visualizing Blocks Per Game Distribution")
```

## Blocks Per Minute Distribution

```{r}
stats_2019 <- stat_2019[order(stat_2019$blkPerMinute),]

#Mean
mean(stats_2019$blkPerMinute)
#Standard Deviation
sd(stats_2019$blkPerMinute)
#Spread
quantile(stats_2019$blkPerMinute)

x <- seq(0, size, 1)
y <- dnorm(stats_2019$blkPerMinute, mean(stats_2019$blkPerMinute), sd(stats_2019$blkPerMinute))

#Graph
ggplot(stats_2019) + 
  geom_line(aes(x = x, y = blkPerMinute), color = "red") +
  xlab("Player") +
  ylab("Steals Per Minute") +
  ggtitle("Visualizing Blocks Per Minute Distribution")

```





